Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-02-09 ribose-paper.RMD
further updates to results

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme)
```

@@

```{r glucose data + model}
# Scripts: gluc.responsechange.R, gluc.fig

# Data
gluc.dat <- read_excel("./data/glucose/fingerstick.xlsx", na = "NA")
gluc.dat$sample_time <- as.character(gluc.dat$sample_time)

# Data handling
glu.dat <- gluc.dat %>%
        mutate(time = if_else(sample_time == "0",
                              "baseline",
                              if_else(sample_time == "45",
                                      "min45",
                                      if_else(sample_time == "90",
                                              "min90",
                                              if_else(sample_time == "120",
                                                      "min120",
                                                      if_else(sample_time == "135",
                                                              "min135",
                                                              if_else(sample_time == "150",
                                                                      "min150",
                                                                      if_else(sample_time == "270",
                                                                              "min270", sample_time))))))),
               time = factor(time, levels = c("baseline", "min45", "min90", "min120", "min135", "min150", "min270"))) %>%
        mutate(glu = as.numeric(glu),
               lak = as.numeric(lak))

# Change data
change_dat.glu <- glu.dat %>%
        dplyr::select(subject, time, glu, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(glu = mean(glu, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = glu) %>%
        #print()

        ungroup() %>%
        mutate(change.45 = log(min45)-log(baseline),
               change.90 = log(min90)-log(baseline),
               change.120 = log(min120)-log(baseline),
               change.135 = log(min135)-log(baseline),
               change.150 = log(min150)-log(baseline),
               change.270 = log(min270)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.45, change.90, change.120, change.135, 
               change.150, change.270) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.45:change.270))

# Model
m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat.glu)
resplot.glu <- plot(m1)
sum.glu <- summary(m1)

# Emmeans
confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
        data.frame()

# For inline plotting
bcoef <- coef(summary(m1))
pscient.120 <- round(bcoef[8,5], 30)
p.120 <- sprintf("%.5f", 1.700971e-16)
pscient.135 <- round(bcoef[9,5], 30)
p.135 <- sprintf("%.5f", 4.299568e-11)
pscient.150 <- round(bcoef[10,5], 30)
p.150 <- sprintf("%.5f", 1.445712e-11)
p.270 <- round(bcoef[11,5], 2)
```

```{r c-peptide data + model}
## Data

ins.dat <- read_excel("./data/glucose/ribose.insulin.xlsx", na = "NA")

code <- read_excel("data/code_key.xlsx", na = "NA") %>%
  select(-leg) %>%
  print()

## Data handling

ins.dat$sample.time <- as.character(ins.dat$sample.time)

ins.dat2 <- ins.dat %>%
  inner_join(code) %>%
  mutate(sample.time = as.character(sample.time)) %>%
  mutate(time = if_else(sample.time == "0",
                        "pre",
                        if_else(sample.time == "0.1",
                                "baseline",
                                if_else(sample.time == "45",
                                "min90",
                                if_else(sample.time == "90",
                                        "min120",
                                        if_else(sample.time == "120",
                                                "min150",
                                                if_else(sample.time == "270",
                                                        "min270", sample.time))))))) %>%
  mutate(time = factor(time, levels = c("pre", "baseline", "min90", "min120", "min150", "min270")),
         c.pep = as.numeric(c.pep)) %>%
  print()

## Baseline analysis
# Comparing pre measurements to post day baseline

base.ins <- ins.dat2 %>%
  filter(time %in% c("pre", "baseline")) %>%
  select(subject, time, supplement, c.pep) %>%
  group_by(supplement) %>%
  pivot_wider(names_from = time, 
              values_from = c.pep) %>%
  print()

ins.ttest <- t.test(base.ins$pre, base.ins$baseline, paired = TRUE)

## Change data

ins.change <- ins.dat2 %>%
  #filter(subject != "102") %>%
  dplyr::select(subject, time, c.pep, supplement) %>%
  group_by(subject, time, supplement) %>%
  summarise(mean.pep = mean(c.pep, na.rm = TRUE)) %>%
  pivot_wider(names_from = time, 
              values_from = mean.pep) %>%
  ungroup() %>%
  mutate(change.90 = min90-baseline,
         change.120 = min120-baseline,
         change.150 = min150-baseline,
         change.270 = min270-baseline,
         baseline = baseline - mean(baseline, na.rm = TRUE),
         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
  pivot_longer(names_to = "time",
               values_to = "change",
               cols = (change.90:change.270)) %>%
  print()

saveRDS(ins.change, "./data/data-gen/glucose/cpep.change.RDS")

ins.lchange <- ins.dat2 %>%
  #filter(subject != "102") %>%
  dplyr::select(subject, time, c.pep, supplement) %>%
  group_by(subject, time, supplement) %>%
  summarise(mean.pep = mean(c.pep, na.rm = TRUE)) %>%
  pivot_wider(names_from = time, 
              values_from = mean.pep) %>%
  ungroup() %>%
  mutate(change.90 = log(min90)-log(baseline),
         change.120 = log(min120)-log(baseline),
         change.150 = log(min150)-log(baseline),
         change.270 = log(min270)-log(baseline),
         baseline = baseline - mean(baseline, na.rm = TRUE),
         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
  pivot_longer(names_to = "time",
               values_to = "change",
               cols = (change.90:change.270)) %>%
  print()

saveRDS(ins.lchange, "./data/data-gen/glucose/cpep.logchange.RDS")

## Linear mixed effects model

m1.1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = ins.lchange) # Main analysis, using log-transformed data due to heteroscedasticity

m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = ins.change) # For emmeans and figure illustration

plot(m1)
plot(m2)

summary(m1.1)

## Emmeans

library(pbkrtest)

# For analysis
insulin.change <- confint.m1.1 <- confint(emmeans(m1.1, specs = ~"supplement|time")) %>%
  data.frame()

saveRDS(insulin.change, "./data/data-gen/glucose/insulin.change.RDS")

# For figure
insulin.figchange <- confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
  data.frame()

saveRDS(insulin.figchange, "./data/data-gen/glucose/insulin.figchange.RDS")

# For inline plotting
cpcoef <- coef(summary(m1.1))
pscient.120 <- round(cpcoef[6,5], 30)
cpp.120 <- sprintf("%.5f", 3.229267e-15)
cpscient.150 <- round(cpcoef[7,5], 30)
cpp.150 <- sprintf("%.5f", 1.414491e-15)
cpp.270 <- round(cpcoef[8,5], 2)

@@

```{r nutrition table}

@@

```{r nutrition analysis}
#summarise macro
#nutritiontable with pro.pr kg bodyweight pr. group
#laod dxa data and take out weight and subject to join with nutrition data
fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
        select(weight, subject)

@@

```{r nutrition change fig}
# Associated scripts:
# - nut.change
# - nut.change.figs

@@

```{r glucose + c-peptide fig}
# Figure
pos <- position_dodge(width = 0.2)
gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")

@@

plot_grid(gluc.fig,
          cpep.fig,
          ncol = 1, nrow = 2)

@@

```{r total RNA data}

code <- read_excel("./data/code.key.xlsx")
rna_dat <- readRDS("./data/data-gen/rna/rna_complete.RDS")

@@

```{r total RNA data handling}

# Scripts
# totrna_analysis
rna_dat2 <- rna_dat %>%
        inner_join(code) %>%
        mutate(time = if_else(time %in% c("T1", "T2"), "pre", "post"), 
               time = factor(time, levels = c("pre", "post"))) %>%
        # filter(outlier == "in") %>%
        group_by(subject, time, rep, outlier, technical, biopsy, supplement) %>%
        summarise(weight = mean(weight), 
                  RNA = mean(RNA_tot)) %>%
        ungroup() %>%
        mutate(weight.mc = weight/mean(weight)) %>%
        mutate(RNA.weight = RNA / weight)

#### Basic model 
change_dat <- rna_dat2 %>%
   #  filter(technical == "in", 
   #        outlier == "in") %>%
    dplyr::select(subject:rep, supplement, RNA.weight) %>%

    group_by(subject, time, supplement) %>%
    summarise(RNA.weight = mean(RNA.weight, na.rm = TRUE)) %>%

    pivot_wider(names_from = time, 
                values_from = RNA.weight) %>%
    ungroup() %>%
    mutate(change = log(post) - log(pre), 
           pre = pre - mean(pre, na.rm = TRUE), 
           supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))

```{r total RNA model}
# Scripts
# totrna_analysis
# Create model: 
m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
           data = change_dat)

resplot.rna <- plot(m1)
sum.rna <- summary(m1)

### Get estimated means from the model
confint.totrna <- confint(emmeans(m1, specs = ~"supplement")) %>%
    data.frame() 
# For inline reporting
trcoef <- coef(summary(m1))
trdiff <- round(trcoef[3,1], 2)
trdiffp <- round(trcoef[3,5], 2)
trdiffse <- round(trcoef[3,2], 2)

```{r rRNA data}
code_key <- read_excel("./data/code.key.xlsx") %>%
        dplyr::select(subject, supplement, time)
qpcr_data  <- readRDS("./data/data-gen/rna/qpcr-data.RDS")
samples <- read_excel("./data/rna/RNA.raw.xlsx") %>%
        dplyr::select(subject, time_rep, sample, weight)

```{r rRNA data handling}
# Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

nf <- qpcr_data %>%
        mutate(sample = as.numeric(sample)) %>% 
        inner_join(samples) %>%
        separate(time_rep, into = c("time","rep"), sep = "rna") %>%
        mutate(rep = paste0("cdna", rep)) %>%
        inner_join(code_key) %>% 
        mutate(time = if_else(time %in% c("T1", "T2"), "Pre", "Post"), 
               time = factor(time, levels = c("Pre", "Post")), 
               expr = eff ^ -cq) %>%

        filter(target %in% c("Lambda F2R2", "Lambda F3R3"), 
               cq < 35) %>%
        separate(target, into = c("target", "nf_primer")) %>%

        mutate(nf.w = expr * weight, 
               nf.w = nf.w / max(nf.w)) %>%

        dplyr::select(sample, subject, time, rep, supplement, nf.w, nf_primer)

qpcr_data2 <- qpcr_data %>%
        mutate(sample = as.numeric(sample)) %>% 
        inner_join(samples) %>%
        inner_join(nf) %>%
        separate(time_rep, into = c("time","rep"), sep = "rna") %>%
        mutate(rep = paste0("cdna", rep)) %>%
        inner_join(code_key) %>% 
        mutate(time = if_else(time %in% c("T1", "T2"), "Pre", "Post"), 
               time = factor(time, levels = c("Pre", "Post")), 
               expr = eff ^ -cq) %>%

        filter(!(target %in% c("Lambda F2R2", "Lambda F3R3"))) %>%   
        inner_join(nf)

saveRDS(qpcr_data2, "./data/data-gen/rna/qpcr_data2.RDS")

# Plotting mean expression per treatment for rRNA
initmod <- qpcr_data2 %>%
        mutate(norm.expr = log(expr / nf.w)) %>%
        group_by(time, supplement, target) %>%
        summarise(m = mean(norm.expr, na.rm = TRUE))  %>%
        #print()

        ggplot(aes(time, m, color = supplement, group = supplement)) + 
        geom_line() +
        geom_point() +
        facet_wrap( ~ target, scales = "free")
qdat <- qpcr_data2 %>%
        mutate(target = gsub("rRNA", "", target),
               target = gsub("  ", " ", target),
               target = paste0("trg_", target)) %>%
        separate(target, into = c("target", "primer"), sep = " ") %>%

        # Create a weight-normalized variable
        mutate(nf.expr = log(expr / nf.w), 
               nf.w = scale(nf.w),
               # Technical random effect
               technical = paste(subject, time, supplement, rep, sep = "_"),
               biological = paste0("S", sample)) %>%
        filter(!(target %in% c("trg_MHC1", "trg_MHC2A", "trg_MHC2X")))

## Change data
rrna18 <- qdat %>%
        filter(target == "trg_18s")
rrna28 <- qdat %>%
        filter(target == "trg_28s")
rrna5.8 <- qdat %>%
        filter(target == "trg_5.8s")
rrna5 <- qdat %>%
        filter(target == "trg_5s")
rrna47 <- qdat %>%
        filter(target == "trg_47s")

# Change scores
# 18S
change.18 <- rrna18 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) 
# 28S
change.28 <- rrna28 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) 
# 5.8S
change.5.8 <- rrna5.8 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
# 5S
change.5 <- rrna5 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
# 47S
change.47 <- rrna47 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
```

```{r rRNA data modelling}
# Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

@@

# 18S
m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.18)
resplot.18 <- plot(m1)
sum.18 <- summary(m1)

# 28S
m2 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.28)
resplot.28 <- plot(m2)
sum.28 <- summary(m2)

# 5.8S
m3 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.5.8)
resplot.58 <- plot(m3)
sum.58 <- summary(m3)

# 5S
m4 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.5)
resplot.5 <- plot(m4)
sum.5 <- summary(m4)

# 47S
m5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.47)
resplot.47 <- plot(m5)
sum.47 <- summary(m5)

@@

# 18S
confint.18 <- confint(emmeans(m1, specs = ~"supplement")) %>%
        data.frame()

# 28S
confint.28 <- confint(emmeans(m2, specs = ~"supplement")) %>%
        data.frame()

# 5.8S
confint.5.8 <- confint(emmeans(m3, specs = ~"supplement")) %>%
        data.frame()

# 5S
confint.5 <- confint(emmeans(m4, specs = ~"supplement")) %>%
        data.frame()

# 47S
confint.47 <- confint(emmeans(m5, specs = ~"supplement")) %>%
        data.frame()

@@

# 18S
rr18coef <- coef(summary(m1))
rr18diffp <- round(rr18coef[3,5], 3)
rr18diffse <- round(rr18coef[3,2], 3)

# 28S
rr28coef <- coef(summary(m2))
rr28diffp <- round(rr28coef[3,5], 3)
rr28diffse <- round(rr28coef[3,2], 3)

# 5.8S
rr58coef <- coef(summary(m3))
rr58diffp <- round(rr58coef[3,5], 3)
rr58diffse <- round(rr58coef[3,2], 3)

# 5S
rr5coef <- coef(summary(m4))
rr5diffp <- round(rr5coef[3,5], 3)
rr5diffse <- round(rr5coef[3,2], 3)

# 47S
rr47coef <- coef(summary(m5))
rr47diffp <- round(rr47coef[3,5], 3)
rr47tdiffp <- round(rr47coef[1,5], 3)
rr47diffse <- round(rr47coef[3,2], 3)

@@

```{r western modelling}
# Script: western.change

@@

prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS")

# Change analysis

#prot.dat %>%
 # group_by(time, supplement, target) %>%
  #mutate(norm.sign = log(tpl/norm.sign)) %>% # Log transforms and normalises
  # norm.sign by tpl
  #summarise(m = mean(norm.sign, na.rm = TRUE)) %>%
  #ggplot(aes(time, m, color = supplement, group = supplement)) +
  #geom_line() +
  #geom_point() +
  #facet_wrap( ~ target, scales = "free")

# Change score

pchange <- prot.dat %>%
  dplyr::select(subject, time, sample.id, target, norm.sign, supplement) %>%
  group_by(subject, time, target, supplement) %>%
  summarise(mns = mean(norm.sign, na.rm = TRUE)) %>%
  pivot_wider(names_from = time,
              values_from = mns) %>%
  ungroup() %>%
  mutate(change = log(post)-log(pre),
         pre = pre - mean(pre, na.rm = TRUE),
         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))

@@

cmyc <- pchange %>%
  filter(target == "cmyc")

ubf <- pchange %>%
  filter(target == "ubf")

rps6 <- pchange %>%
  filter(target == "rps6")

@@

```{r training data}
lp.load <- read_excel("./data/training/ribose_training.xlsx", na = "NA") %>%
        select(subject, timepoint, supplement, exercise, load) %>%
        filter(exercise != "KE")

lp.rm <- read_excel("./data/tests/ribose.1rm.xlsx", na = "NA") %>%
        select(subject, supplement, exercise, rm) %>%
        filter(exercise != "KE")

tot.load <- read_excel("./data/training/ribose_training.xlsx", na = "NA") %>%
        select(subject, timepoint, supplement, exercise, load)

tot.rm <- read_excel("./data/tests/ribose.1rm.xlsx", na = "NA") %>%
        select(subject, supplement, exercise, rm)

tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
        select(subject, timepoint, tot.volume, supplement)

@@

```{r training data handling}
# Scripts
# tot.vol.change
# Trainingint.linear
# rpe-analysis

>>>>

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest)
```

@@

```{r glucose data + model}
## Scripts:
# - gluc.responsechange.R

# Data
glu.change <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Untransformed data, used in figures
glu.logchange <- readRDS("./data/data-gen/glucose/glu.logchange.RDS") # Log-transformed data, used in the analysis

## Model
# Untransformed
glu.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = glu.change)
glu.res <- plot(glu.mod)
glu.sum <- summary(glu.mod)

# Log-transformed
glu.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = glu.logchange)
glu.lres <- plot(glu.lmod)
glu.lsum <- summary(glu.lmod)

## Emmeans
# Untransformed
glu.emm <- confint(emmeans(glu.mod, specs = ~"supplement|time")) %>%
        data.frame()

# Log-transformed
glu.lemm <- confint(emmeans(glu.lmod, specs = ~"supplement|time")) %>%
        data.frame()

# For inline plotting
bcoef <- coef(summary(glu.lmod))
pscient.120 <- round(bcoef[8,5], 30)
p.120 <- sprintf("%.5f", 1.700971e-16)
pscient.135 <- round(bcoef[9,5], 30)
p.135 <- sprintf("%.5f", 4.299568e-11)
pscient.150 <- round(bcoef[10,5], 30)
p.150 <- sprintf("%.5f", 1.445712e-11)
p.270 <- round(bcoef[11,5], 2)
```

```{r c-peptide data + model}
## Scripts
# - insulin.change.R

## Data
cpep.change <- readRDS("./data/data-gen/glucose/cpep.change.RDS") # Untransformed data, for figures
cpep.lchange <- readRDS("./data/data-gen/glucose/cpep.logchange.RDS") # Log-transformed data, for analysis

## Linear mixed effects model
# Untransformed
cpep.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = cpep.change) 
cpep.res <- plot(cpep.mod)
cpep.sum <- summary(cpep.mod)

# Log-transformed
cpep.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = cpep.lchange) 
cpep.lres <- plot(cpep.lmod)
cpep.lsum <- summary(cpep.lmod)

## Emmeans
# Untransformed
cpep.emm <- confint(emmeans(cpep.mod, specs = ~"supplement|time")) %>%
  data.frame()

# Log-transformed
cpep.lemm <- confint(emmeans(cpep.lmod, specs = ~"supplement|time")) %>%
  data.frame()

# For inline plotting
cpcoef <- coef(summary(cpep.lmod))
pscient.120 <- round(cpcoef[6,5], 30)
cpp.120 <- sprintf("%.5f", 3.229267e-15)
cpscient.150 <- round(cpcoef[7,5], 30)
cpp.150 <- sprintf("%.5f", 1.414491e-15)
cpp.270 <- round(cpcoef[8,5], 2)

@@

```{r nutrition table}
## Scripts
# - nut.table.R
# - nut.analysis.R
# - dxa.R

@@

```{r nutrition analysis}
## Scripts
# - nut.analysis.R
# - dxa.R

#summarise macro
#nutritiontable with pro.pr kg bodyweight pr. group
#laod dxa data and take out weight and subject to join with nutrition data
fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
        select(weight, subject)

@@

```{r nutrition change analysis and fig}
# Associated scripts:
# - nut.change
# - nut.change.figs

@@

```{r glucose + c-peptide fig}
## Scripts
# - gluc.fig.R
# - cpep.figs.R

# Separate figures
pos <- position_dodge(width = 0.2)

gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")

@@

# Joined figure
plot_grid(gluc.fig,
          cpep.fig,
          ncol = 1, nrow = 2)

@@

```{r total RNA model}
## Scripts
# - totrna.cleanup.R
# - totrna.analysis.R

# Data
trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")

# Create model: 
trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
           data = trna.lchange)

trna.lres <- plot(trna.lmod)
trna.lsum <- summary(trna.lmod)

### Get estimated means from the model
trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
    data.frame()

# For inline reporting
trcoef <- coef(summary(trna.lmod))
trdiff <- round(trcoef[3,1], 2)
trdiffp <- round(trcoef[3,5], 2)
trdiffse <- round(trcoef[3,2], 2)

@@

```{r rRNA data model}
## Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

# Log-transformed data
lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")

@@

# 18S
lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.18)
lres.18 <- plot(lmod.18)
lsum.18 <- summary(lmod.18)

# 28S
lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.28)
lres.28 <- plot(lmod.28)
lsum.28 <- summary(lmod.28)

# 5.8S
lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.58)
lres.58 <- plot(lmod.58)
lsum.58 <- summary(lmod.58)

# 5S
lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.5)
lres.5 <- plot(lmod.5)
lsum.5 <- summary(lmod.5)

# 47S
lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = lchange.47)
lres.47 <- plot(lmod.47)
lsum.47 <- summary(lmod.47)

@@

# 18S
lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
        data.frame()

# 28S
lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
        data.frame()

# 5.8S
lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
        data.frame()

# 5S
lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
        data.frame()

# 47S
lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
        data.frame()

@@

# 18S
rr18coef <- coef(summary(lmod.18))
rr18diffp <- round(rr18coef[3,5], 3)
rr18diffse <- round(rr18coef[3,2], 3)

# 28S
rr28coef <- coef(summary(lmod.28))
rr28diffp <- round(rr28coef[3,5], 3)
rr28diffse <- round(rr28coef[3,2], 3)

# 5.8S
rr58coef <- coef(summary(lmod.58))
rr58diffp <- round(rr58coef[3,5], 3)
rr58diffse <- round(rr58coef[3,2], 3)

# 5S
rr5coef <- coef(summary(lmod.5))
rr5diffp <- round(rr5coef[3,5], 3)
rr5diffse <- round(rr5coef[3,2], 3)

# 47S
rr47coef <- coef(summary(lmod.47))
rr47diffp <- round(rr47coef[3,5], 3)
rr47tdiffp <- round(rr47coef[1,5], 3)
rr47diffse <- round(rr47coef[3,2], 3)

@@

```{r western modelling}
## Script: 
# - western.import.R
# - western.change.R

@@

prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data

@@

cmyc <- prot.change %>%
  filter(target == "cmyc")

ubf <- prot.change %>%
  filter(target == "ubf")

rps6 <- prot.change %>%
  filter(target == "rps6")

@@

```{r training data model}
## Scripts
# - volume.change.R
# - ss.analysis.R
# - ss.cleanup.R
# - trainint.analysis.R
# - trainint.cleanup.R

# Log-transformed data
vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS")
ss.lchange <- readRDS("./data/data-gen/training/ss.lchange.RDS")
int.lchange <- readRDS("./data/data-gen/training/int.lchange.RDS")

### Training volume
# Fold-change model
vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = vol.lchange)
vol.lres <- plot(vol.lmod)
vol.lsum <- summary(vol.lmod)

# Fold-change estimated means
vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
        data.frame()

### Session score (RPE)
# Model
ss.lmod <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
                     data = ss.lchange)
ss.lres <- plot(ss.lmod)
ss.lsum <- summary(ss.lmod)

# Session score emmeans
ss.lemm <- confint(emmeans(ss.lmod, specs = ~"supplement|time")) %>%
        data.frame()

### Training intensity
int.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = int.lchange)
int.lres <- plot(int.lmod)
int.lsum <- summary(int.lmod)

# Emmeans
int.lemm <- confint(emmeans(int.lmod, specs = ~"supplement|time")) %>%
        data.frame()

## Training volume
# Diff 4
vcoef <- coef(summary(vol.lmod))
vdiff.4 <- round(vcoef[9,1], 3)
vdiffp.4 <- round(vcoef[9,5], 3)
vdiffse.4 <- round(vcoef[9,2], 3)
# Diff 5
vdiff.5 <- round(vcoef[10,1], 3)
vdiffp.5 <- round(vcoef[10,5], 3)
vdiffse.5 <- round(vcoef[10,2], 3)
# Diff 6
vdiff.6 <- round(vcoef[11,1], 3)
vdiffp.6 <- round(vcoef[11,5], 3)
vdiffse.6 <- round(vcoef[11,2], 3)
## RPE
scoef <- coef(summary(ss.lmod))
sdiffp <- round(scoef[11, 5], 3)
sdiffse <- round(scoef[11, 2], 3)

# Diff time 2
icoef <- coef(summary(int.lmod))
idiff2 <- round(icoef[7,1], 3)
idiffp2 <- round(icoef[7,5], 3)
idiffse2 <- round(icoef[7,2], 3)
# Diff time 3
idiff3 <- round(icoef[8,1], 3)
idiffp3 <- round(icoef[8,5], 3)
idiffse3 <- round(icoef[8,2], 3)
# Diff time 4
idiff4 <- round(icoef[9,1], 3)
idiffp4 <- round(icoef[9,5], 3)
idiffse4 <- round(icoef[9,2], 3)
# Diff time 5
idiff5 <- round(icoef[10,1], 3)
idiffp5 <- round(icoef[10,5], 3)
idiffse5 <- round(icoef[10,2], 3)
# Diff time 6
idiff6 <- round(icoef[11,1], 3)
idiffp6 <- round(icoef[11,5], 3)
idiffse6 <- round(icoef[11,2], 3)
```

```{r humac post 5th model}
## Scripts
# - humac.cleanup.R
# - humac.main.analysis.R - this is data gathered at baseline, during the intervention, and after session 5

# Log-transformed data

isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS")
iso60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS")
iso240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS")

# Isometric
isom.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = isom.lchange)
isom.lres <- plot(isom.lmod)
isom.lsum <- summary(isom.lmod)

# Isok.60
iso60.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = iso60.lchange)
iso60.lres <- plot(iso60.lmod)
iso60.lsum <- summary(iso60.lmod)

# Isok.240
iso240.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = iso240.lchange)
iso240.lres <- plot(iso240.lmod)
iso240.lsum <- summary(iso240.lmod)

## Emmeans
# Isometric
isom.lemm <- confint(emmeans(isom.lmod, specs = ~"supplement|time")) %>%
        data.frame()

# Isok.60
iso60.lemm <- confint(emmeans(iso60.lmod, specs = ~"supplement|time")) %>%
        data.frame()

# Isok.240
iso240.lemm <- confint(emmeans(iso240.lmod, specs = ~"supplement|time")) %>%
        data.frame()

# For reporting
# Isom diff
isom5coef <- coef(summary(isom.lmod))
isom5diffp <- round(isom5coef[7, 5], 3)
isom5diffse <- round(isom5coef[7, 2], 3)

# Isok 60 diff
isok605coef <- coef(summary(iso60.lmod))
isok605diffp <- round(isok605coef[7, 5], 3)
isok605diffse <- round(isok605coef[7, 2], 3)

# Isok 260
isok2405coef <- coef(summary(iso240.lmod))
isok2405diffp <- round(isok2405coef[7, 5], 3)
isok2405diffse <- round(isok2405coef[7, 2], 3)

@@

```{r post 6th model}
## Scripts
# - humac.acute.analysis.R - this is data gatherede before and after one RT session (sixth session)
# - humac.cleanup.R

# Data
isom.lchange.ac <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS")
iso60.lchange.ac <- readRDS("./data/data-gen/humac/iso60.lchange.ac.RDS")
iso240.lchange.ac <- readRDS("./data/data-gen/humac/iso240.lchange.ac.RDS")
